### 2020-05-29 writeup/proof_examples.tex

Combining \eqref{hatw_add1} and \eqref{boundv-w}, we obtain that whp,
\be\label{hatv_add1} 
|\hat v - 1|= \OO\left(\cal E\right), \quad \cal E:=\frac{d^2}{\kappa^2} + \frac{\sigma^2}{p\kappa^2} + p^{-1/2+\e} 
\ee
Inserting it into \eqref{eq_te_mtl_2task} and using a similar concentration result as in \eqref{boundv-w}, we get whp,
\begin{align}
\te(\hat{\beta}_t^{\MTL})&=(1+\OO(\cal E))\cdot \left[d^2 + \OO\left(\cal E^2 \kappa^2\right)\right] \tr\left[(X_1^{\top}X_1 +X_2^{\top}X_2)^{-2} (X_1^{\top}X_1)^2\right]\nonumber\\ 
&+(1+\OO(\cal E))\cdot \sigma^2  \bigtr{(X_1^{\top}X_1  + X_2^{\top}X_2)^{-1} }. \nonumber
\end{align}
In order to study the phenomenon of bias-variance trade-off, we need the bias term with $d^2$ and the variance term with $\sigma^2$ to be of the same order. With estimate \eqref{eq_isometric}, we see that 
$$\tr\left[(X_1^{\top}X_1 +X_2^{\top}X_2)^{-2} (X_1^{\top}X_1)^2\right] \sim p,\quad \bigtr{(X_1^{\top}X_1  + X_2^{\top}X_2)^{-1} } \sim \frac{p}{n_1+n_2}.$$
Hence we need to choose that $p\cdot d^2 \sim \sigma^2$. On the other hand, we want the error term $\cal E^2 \kappa^2$ to be much smaller than $d^2$, which leads to the condition $p^{-1+2\e}\kappa^2  \ll d^2 \ll \kappa^2$. The above considerations lead to the following choices of parameters: there exists a constant $c>0$ such that
\be\label{choiceofpara}
pd^2 \sim \sigma^2 \sim 1,\quad p^{-1+c}\kappa^2  \le d^2 \le p^c \kappa^2.
\ee
Under this choice, we can write 
\be\label{simple1}
\begin{split}
\te(\hat{\beta}_t^{\MTL})&=(1+\OO(n^{-\e}))\cdot d^2 \tr\left[(X_1^{\top}X_1 +X_2^{\top}X_2)^{-2} (X_1^{\top}X_1)^2\right] \\ 
&+(1+\OO(n^{-\e}))\cdot \sigma^2  \bigtr{(X_1^{\top}X_1  + X_2^{\top}X_2)^{-1} }  
\end{split}
\ee
whp for some constant $\e>0$.

@@

 Finally, we prove Proposition \ref{prop_covariate}, which shows that $\te(\hat{\beta}^{\MTL})$ is minimized approximately when $M$ is a scalar matrix if there are sufficiently many source datas.

\begin{proof}[Proof of Proposition \ref{prop_covariate}]
........ Let 
$$M_0:=\argmin_{M\in \cal S_{\mu}}(\te(\hat{\beta}_t^{\MTL}))(M).$$ 
We now calculate $(\te(\hat{\beta}_t^{\MTL}))(M_0)$. In this case, as in Lemma \ref{lem_hat_v} we also have that 
\be\label{hatvM}|\hat v-1|=\OO(p^{-1}) \ee 
in the setting Proposition \ref{prop_covariate}. In fact, Lemma \ref{lem_hat_v} was proved assuming that $M=\id$, but its proof can be easily extended to the case with general $M\in \cal S_{\mu}$ by using that $\lambda(M)\in [\mu_{\min},\mu_{\max}]$. We omit the details here.

Now using \eqref{hatvM}, Lemma \ref{lem_cov_shift} and Lemma \ref{prop_model_shift_tight}, we get that whp,
\begin{align*}
(\te(\hat{\beta}_t^{\MTL}))(M_0)= \frac{\sigma^2}{c_1+c_2}\cdot \frac1p\tr\left( \frac{1}{a_1M_0^\top M_0 + a_2}\right)  + \Delta_\beta(M_0),
\end{align*}
where  $ \Delta_{\beta}(M_0)$ satisfies 
\be \nonumber
\begin{split}
&\left| \Delta_{\beta}(M_0) - \frac{d^2\cdot c_1^2}{(c_1 + c_2)^2} \tr\left[{M_0} \frac{(1 + a_3)\id + a_4 {M_0}^{\top}{M_0}}{(a_2 + a_1 {M_0}^{\top}{M_0})^2} {M_0}^{\top}\right]\right| \\
&\le \left(\bigbrace{1 + \sqrt{\frac{1}{c_1}}}^4-1\right)\left( \frac{c_1\mu_{\max}}{(\sqrt{c_1}-1)^2 \mu_{\min} + (\sqrt{c_2}-1)^2}\right)^2 \cdot d^2\tr\left(\Sigma_1\right). 
\end{split}
\ee
From equation \eqref{eq_a2}, we get
$$a_1\ge \frac{c_1-1}{c_1+c_2},\quad a_2\le \frac{c_2}{c_1+c_2}.$$
Then solving equations \eqref{eq_a3} and \eqref{eq_a4}, we get that
\begin{align*}
0\le a_3&= \frac{ \frac{1}{n_2}\sum_{i=1}^p \frac{a_2^2 }{ (  a_2 + \lambda_i^2a_1)^2  }\left(1 - \frac{1}{n_1} \sum_{i=1}^p \frac{\lambda_i^4 a_1^2  }{  (a_2 + \lambda_i^2a_1)^2  }\right) +\left( \frac{1}{n_1}\sum_{i=1}^p \frac{\lambda_i^2 a_1^2}{  (a_2 + \lambda_i^2a_1)^2  }\right) \left(\frac{1}{n_2}\sum_{i=1}^p \frac{  \lambda_i^2 a_2^2 }{ (  a_2 + \lambda_i^2a_1)^2  }\right) } {\left(1- \frac1{n_2}\sum_{i=1}^p \frac{a_2^2}{ (a_2 + \lambda_i^2a_1)^2  }\right) \left(1 - \frac{1}{n_1} \sum_{i=1}^p \frac{\lambda_i^4 a_1^2  }{  (a_2 + \lambda_i^2a_1)^2  }\right) - \left(\frac{1}{n_2}\sum_{i=1}^p \frac{  \lambda_i^2 a_2^2 }{ (  a_2 + \lambda_i^2a_1)^2  }\right) \left(\frac{1}{n_1}\sum_{i=1}^p \frac{\lambda_i^2 a_1^2 }{  (a_2 + \lambda_i^2a_1)^2  }\right)} \\
&\le \frac{c_2^{-1}}{1-c_1^{-1}-c_2^{-1}}
\end{align*}
and
\begin{align*}
0\le a_4&= \frac{\frac{1}{n_1}\sum_{i=1}^p \frac{\lambda_i^2 a_1^2}{  (a_2 + \lambda_i^2a_1)^2  }}{\left(1- \frac1{n_2}\sum_{i=1}^p \frac{a_2^2}{ (a_2 + \lambda_i^2a_1)^2  }\right) \left(1 - \frac{1}{n_1} \sum_{i=1}^p \frac{\lambda_i^4 a_1^2  }{  (a_2 + \lambda_i^2a_1)^2  }\right) - \left(\frac{1}{n_2}\sum_{i=1}^p \frac{  \lambda_i^2 a_2^2 }{ (  a_2 + \lambda_i^2a_1)^2  }\right) \left(\frac{1}{n_1}\sum_{i=1}^p \frac{\lambda_i^2 a_1^2 }{  (a_2 + \lambda_i^2a_1)^2  }\right)} \\
&\le  \frac{c_1^{-1}\cdot \mu_{\min}^{-2}}{1-c_1^{-1}-c_2^{-1}},
\end{align*}
where we also used that 
$$\left(\sum_{i=1}^p \frac{\lambda_i^2  }{  (a_2 + \lambda_i^2a_1)^2  }\right)^2 \le \sum_{i=1}^p \frac{\lambda_i^4 }{  (a_2 + \lambda_i^2a_1)^2  }\cdot \sum_{i=1}^p \frac{1}{  (a_2 + \lambda_i^2a_1)^2  }$$
by Cauchy-Schwarz inequality.

Combining the above estimates, we get that 
\begin{align}\label{approximateteM}
(\te(\hat{\beta}_t^{\MTL}))(M_0) =  \te(M_0) + \cal E,
\end{align}
where
$$\te(M_0):=\frac{\sigma^2}{c_1+c_2}\cdot \frac1p\tr\left( \frac{1}{a_1M_0^\top M_0}\right)  +\frac{d^2\cdot c_1^2}{(c_1 + c_2)^2} \tr\left(\frac{1 + a_3}{a_1^2 M_0^{\top}{M_0}} \right),$$
and the error satisfies
$$|\cal E|\le C\left( \frac{c_2}{c_1} +c_1^{-1/2}\right)\te(M_0),$$
Here the constant $C>0$ depends only on $\mu_{\max}$, $\mu_{\min}$ and $\|\Sigma_1\|$, but otherwise does not depend on $c_1$ and $c_2$.

%With these two bounds, we can easily conclude \eqref{approxteM}. 
%
%We have that the test error satisfies
%\be\label{approxteM}  te(M)\left(1 -  \frac{n_2}{n_1-p} \frac{1}{\lambda_p^2 + \frac{n_2}{n_1-p}}\right)  \le  \frac{\sigma^2}{n_1+n_2}\tr\left( \frac{1}{a_1M^\top M + a_2}\right) \le te(M),\ee
%where $\lambda_p$ is the smallest singular value of $p$ and
%$$te(M):= \frac{\sigma^2}{a_1(n_1+n_2)}\tr\left( \frac{1}{M^\top M}\right) .$$
%Moreover, for all $M$ satisfying \eqref{GMcons}, the minimum of $te(M)$ is attained when $M= a\id$.

Finally using AM-GM inequality, we observe that 
$$\tr\left( \frac{1}{M^\top M}\right) = \sum_{i=1}^p\frac{1}{\lambda_i}$$
is minimized when $\lambda_1 = \cdots\lambda_p=\mu$ under the restriction $\prod_{i=1}^p\lambda_i\le \mu^p$. Hence we get that 
$$ \te (M_0)\le \te(\mu \id).$$
Together with \eqref{approximateteM}, we conclude \eqref{formular_covariate0}.
%, we conclude that the sum $\sum_{i=1}^p\lambda_i^{-1}$ is smallest when $\lambda_1=\cdots=\lambda_p = a$.
\end{proof}

\begin{proof}[Theoretical justification of Example \ref{ex_covariate}]
In the setting of Example \ref{ex_covariate}, equations in \eqref{eq_a2} become
\be\label{compleeq} a_1 + a_2 = 1 - \frac{p}{n_1 + n_2},  \ \ a_1 + \frac{p}{2(n_1 + n_2)}\cdot \bigbrace{\frac{a_1}{a_1 + \lambda^2 a_2} + \frac{a_1}{a_1 + \frac{a_2}{\lambda^2}}} = \frac{n_1}{n_1 + n_2}. \ee
It's not hard to verify that there is only one valid solution $(a_1,a_2)$ to \eqref{compleeq}. After solving these, we get the test error for the target task as follows.
\be\label{testcomple}
te(\lambda)=\frac{p}{2(n_1 + n_2)} \cdot \bigbrace{\frac{1}{\frac{a_1}{\lambda^2} + a_2} + \frac{1}{a_1\lambda^2 + a_2}}.\ee

First we notice that the curves in Figure \ref{fig_te_complement} all cross at the point $n_1=n_2$. In fact, if $n_1=n_2$, then it is easy to observe that $a_1=a_2=(1-\gamma)/2$ is the solution to equation \eqref{compleeq}, where we denote $ \gamma=p/(n_1+n_2)$. Then for any $\lambda$, the test error in \eqref{testcomple} takes the value
$$te(\lambda)= \frac{\gamma}{2}\frac{1}{(1-\gamma)/2}=\frac{p}{n_1+n_2-p}.$$

%Second, from Figure \ref{fig_te_complement} we observe that the complementary cases with $\lambda>1$ is better than the case without covariate shift (i.e. $M=\id$ case) when $n_1<n_2$. On the other hand, if we have enough source task data such that $n_1>n_2$, then it is always better to have no covariate shift.
This phenomenon can be also explained using our theory. With \eqref{compleeq}, we can write
$$te(\lambda)=\frac{\gamma}{2} \cdot \bigbrace{\frac{1}{\frac{a_1}{\lambda^2} + (1-\gamma - a_1)} + \frac{1}{a_1\lambda^2 + (1-\gamma - a_1)}}.$$
We can compute that
\begin{align*}
te(\lambda) - te(1)&= \frac{\gamma}{2(1-\gamma)} (\lambda^2-1)a_1\cdot \bigbrace{  \frac{1}{ -a_1(\lambda^2-1)+(1-\gamma)\lambda^2 } - \frac{1}{a_1(\lambda^2-1) + (1-\gamma)}} \\
&= \frac{\gamma}{2(1-\gamma)} (\lambda^2-1)^2 a_1\cdot  \frac{2a_1 - (1-\gamma) }{[-a_1(\lambda^2-1)+(1-\gamma)\lambda^2 ][a_1(\lambda^2-1) + (1-\gamma)]} .
\end{align*}
If $n_1>n_2$, we have $a_1>(1-\gamma)/2$ (because $a_1>a_2$ as observed from the equation \eqref{compleeq}), and hence $te(\lambda)>te(1)$. Otherwise if $n_1< n_2$, we have $a_1< (1-\gamma)/2$, and hence $te(\lambda)< te(1)$.
\end{proof}

@@

\subsection{Proofs for Section \ref{sec_data_size}}\label{app_proof_data}

With similar techniques, we then show Proposition \ref{prop_var_transition}, which gives a transition threshold with respect to the difference between the noise levels of the two tasks.

>>>>

Combining \eqref{hatw_add1} and \eqref{boundv-w}, we obtain that whp,
\be\label{hatv_add1} 
|\hat v - 1|= \OO\left(\cal E\right), \quad \cal E:=\frac{d^2}{\kappa^2} + \frac{\sigma^2}{p\kappa^2} + p^{-1/2+\e} 
\ee
Inserting it into \eqref{eq_te_mtl_2task} and using a similar concentration result as in \eqref{boundv-w}, we get whp,
\begin{align}
\te(\hat{\beta}_t^{\MTL})&=(1+\OO(\cal E))\cdot \left[d^2 + \OO\left(\cal E^2 \kappa^2\right)\right] \tr\left[(X_1^{\top}X_1 +X_2^{\top}X_2)^{-2} (X_1^{\top}X_1)^2\right]\nonumber\\ 
&+(1+\OO(\cal E))\cdot \sigma^2  \bigtr{(X_1^{\top}X_1  + X_2^{\top}X_2)^{-1} }. \nonumber
\end{align}
In order to study the phenomenon of bias-variance trade-off, we need the bias term with $d^2$ and the variance term with $\sigma^2$ to be of the same order. With estimate \eqref{eq_isometric}, we see that 
$$\tr\left[(X_1^{\top}X_1 +X_2^{\top}X_2)^{-2} (X_1^{\top}X_1)^2\right] \sim p,\quad \bigtr{(X_1^{\top}X_1  + X_2^{\top}X_2)^{-1} } \sim \frac{p}{n_1+n_2}.$$
Hence we need to choose that $p\cdot d^2 \sim \sigma^2$. On the other hand, we want the error term $\cal E^2 \kappa^2$ to be much smaller than $d^2$, which leads to the condition $p^{-1+2\e}\kappa^2  \ll d^2 \ll \kappa^2$. The above considerations lead to the following choices of parameters: there exists a constant $c>0$ such that
\be\label{choiceofpara}
pd^2 \sim \sigma^2 \sim 1,\quad p^{c}\kappa^2  \le \sigma^2 \le p^{1-c} \kappa^2.
\ee
Under this choice, we can write 
\be\label{simple1}
\begin{split}
\te(\hat{\beta}_t^{\MTL})&=(1+\OO(n^{-\e}))\cdot d^2 \tr\left[(X_1^{\top}X_1 +X_2^{\top}X_2)^{-2} (X_1^{\top}X_1)^2\right] \\ 
&+(1+\OO(n^{-\e}))\cdot \sigma^2  \bigtr{(X_1^{\top}X_1  + X_2^{\top}X_2)^{-1} }  
\end{split}
\ee
whp for some constant $\e>0$.

@@

 Finally, we prove Proposition \ref{prop_covariate}, which shows that $\te(\hat{\beta}^{\MTL})$ is minimized approximately when $M$ is a scalar matrix where there is no model bias and enough source data.

\begin{proof}[Proof of Proposition \ref{prop_covariate}]
Let 
$$M_0:=\argmin_{M\in \cal S_{\mu}}g(M).$$ 
We now calculate $g(M_0)$. With the same arguments as in Lemma \ref{lem_hat_v} we can show that \eqref{hatv_add1} holds. Moreover, if the parameters are chosen such that $p^{c}\kappa^2  \le \sigma^2 \le p^{1-c} \kappa^2$ (recall \eqref{choiceofpara}), we can simplify
\be \nonumber
\begin{split}
g(M_0)&=(1+\OO(p^{-\e}))\cdot \sigma^2  \bigtr{\Sigma_2(X_1^{\top}X_1  + X_2^{\top}X_2)^{-1} }  ,
\end{split}
\ee
with high probability for any constant $\e>0$. In fact, Lemma \ref{lem_hat_v} was proved assuming that $M=\id$, but its proof can be easily extended to the case with general $M\in \cal S_{\mu}$ by using that $\lambda(M)\in [\mu_{\min},\mu_{\max}]$. We omit the details here.

Now using Lemma \ref{lem_cov_shift}, we can obtain that with high probability,
\begin{align}\label{gvar_extra}
g(M_0)= \frac{\sigma^2}{\rho_1+\rho_2}\cdot \frac1p\tr\left( \frac{1}{a_1(M_0)\cdot M_0^\top M_0 + a_2(M_0)}\right) \cdot \left(1 +\OO(p^{-\e})\right).
\end{align}
From equation \eqref{eq_a12extra}, it is easy to obtain the following estimates on $ a_1(M)$ and $a_2(M)$ for any $M\in \cal S_\mu$:
\be\label{est_a12extra}
\frac{\rho_1-1}{\rho_1+\rho_2} < a_1(M)<  \frac{\rho_1+\rho_2-1}{\rho_1+\rho_2},\quad a_2(M) < \frac{\rho_2}{\rho_1+\rho_2}.
\ee
Inserting \eqref{est_a12extra} into \eqref{gvar_extra} and using $\lambda(M_0^\top M_0)\ge \mu_{\min}^2$, we obtain that with high probability,
\begin{align}\label{approximateteM}
\left(1+\frac{\rho_2}{(\rho_1-1)\mu_{\min}^2}\right)^{-1}\wt g(M_0) \cdot \left(1 - \OO(p^{-\e})\right) \le g(M_0) \le \wt g(M_0) \cdot \left(1 +\OO(p^{-\e})\right),
\end{align}
where
$$\wt g(M_0):=\frac{\sigma^2}{(\rho_1+\rho_2)a_1(M_0)}\cdot \frac1p\tr\left( \frac{1}{M_0^\top M_0}\right) .$$
%With these two bounds, we can easily conclude \eqref{approxteM}. 
%
%We have that the test error satisfies
%\be\label{approxteM}  te(M)\left(1 -  \frac{n_2}{n_1-p} \frac{1}{\lambda_p^2 + \frac{n_2}{n_1-p}}\right)  \le  \frac{\sigma^2}{n_1+n_2}\tr\left( \frac{1}{a_1M^\top M + a_2}\right) \le te(M),\ee
%where $\lambda_p$ is the smallest singular value of $p$ and
%$$te(M):= \frac{\sigma^2}{a_1(n_1+n_2)}\tr\left( \frac{1}{M^\top M}\right) .$$
%Moreover, for all $M$ satisfying \eqref{GMcons}, the minimum of $te(M)$ is attained when $M= a\id$.
By AM-GM inequality, we observe that 
$$\tr\left( \frac{1}{M^\top M}\right) = \sum_{i=1}^p\frac{1}{\lambda_i}$$
is minimized when $\lambda_1 = \cdots\lambda_p=\mu$ under the restriction $\prod_{i=1}^p\lambda_i\le \mu^p$. Hence we get that 
\be\label{AMGM}\wt g(M_0) \le \frac{\sigma^2}{\mu^2 (\rho_1+\rho_2)a_1(M_0)}.\ee

On the other hand, when $M=\mu \id$, applying Lemma \ref{lem_cov_shift} we can obtain that with high probability,
\begin{align}\label{gvar_extra2}
\begin{split}
g(\mu \id)&= \frac{\sigma^2}{\rho_1+\rho_2}\cdot \frac1p\tr\left( \frac{1}{\mu^2 a_1 (\mu\id) + a_2(\mu\id)}\right) \cdot \left(1 +\OO(p^{-\e})\right)\\
&\le \frac{\sigma^2}{\mu^2(\rho_1+\rho_2)a_1 (\mu\id)}.
\end{split}
\end{align}
Combining \eqref{est_a12extra}, \eqref{approximateteM}, \eqref{AMGM} and \eqref{gvar_extra2}, we conclude the proof.
%, we conclude that the sum $\sum_{i=1}^p\lambda_i^{-1}$ is smallest when $\lambda_1=\cdots=\lambda_p = a$.
\end{proof}

%--------old arguments
%With the same arguments as in Lemma \ref{lem_hat_v} we can show that \eqref{hatv_add1} holds, and under the choice of parameters \eqref{choiceofpara},
%\be \nonumber
%\begin{split}
%g(M)&=(1+\OO(n^{-\e}))\cdot d^2 \tr\left[(X_1^{\top}X_1)(X_1^{\top}X_1 +X_2^{\top}X_2)^{-1}\Sigma_2(X_1^{\top}X_1 +X_2^{\top}X_2)^{-2} (X_1^{\top}X_1)\right] \\ 
%&+(1+\OO(n^{-\e}))\cdot \sigma^2  \bigtr{\Sigma_2(X_1^{\top}X_1  + X_2^{\top}X_2)^{-1} }  =: \delta_\beta(M)+ g_{\vari}(M),
%\end{split}
%\ee
%with high probability for any constant $\e>0$. In fact, Lemma \ref{lem_hat_v} was proved assuming that $M=\id$, but its proof can be easily extended to the case with general $M\in \cal S_{\mu}$ by using that $\lambda(M)\in [\mu_{\min},\mu_{\max}]$. We omit the details here. 
%
%
%With Lemma \ref{lem_cov_shift}, we can obtain that with high probability,
%\begin{align}\label{gvar_extra}
%g_{\vari}(M_0)= \frac{\sigma^2}{\rho_1+\rho_2}\cdot \frac1p\tr\left( \frac{1}{a_1M_0^\top M_0 + a_2}\right)  +\OO(p^{-1/2+\e}),
%\end{align}
%where $(a_1,a_2)$ satisfies \eqref{eq_a12extra}. On the other hand, by Theorem \ref{thm_model_shift} we can estimate $\delta_\beta$ as
%\be \label{deltabeta_extra}
%\begin{split}
%&\left| \delta_{\beta}(M_0) - \frac{d^2\cdot \rho_1^2}{(\rho_1 + \rho_2)^2} \tr\left[\Sigma_1^{1/2}{M_0} \frac{(1 + a_3)\id + a_4 {M_0}^{\top}{M_0}}{(a_2 + a_1 {M_0}^{\top}{M_0})^2} {M_0}^{\top}\Sigma_1^{1/2}\right]\right| \le \delta + \OO(p^{1-\e}d^2),
%\end{split}
%\ee
%where $(a_3,a_4)$ satisfies \eqref{eq_a34extra} and $\delta$ is defined in \eqref{eq_deltaextra}.
%
%
%Next we derive some estimates for $a_i$, $i=1,2,3,4$. From equation \eqref{eq_a12extra}, it is easy to see that
%\be\label{est_a12extra}a_1\ge \frac{\rho_1-1}{\rho_1+\rho_2},\quad a_2\le \frac{\rho_2}{\rho_1+\rho_2}.\ee
%Then solving \eqref{eq_a34extra}, we get that
%\begin{align}
%0\le a_3&= \frac{ b_0 (\rho_1a_1^{-2}-b_2) + b_1^2 }{  \left(\rho_2  a_2^{-2}-  b_0\right)\left(\rho_1 a_1^{-2} -  b_2  \right) - b_1^2} \le \frac{ b_0 \rho_1a_1^{-2}  }{  \rho_1\rho_2  a_1^{-2}a_2^{-2}-  b_0\rho_1 a_1^{-2} - b_2\rho_2  a_2^{-2}}\nonumber\\
%&\le \frac{\rho_1}{\rho_1\rho_2-\rho_1 -\rho_2 },\label{est_a3extra}
%\end{align}
%and
%\begin{align}
%0\le a_4&= \frac{b_1\rho_2 a_2^{-2}}{\left(\rho_2  a_2^{-2}-  b_0\right)\left(\rho_1 a_1^{-2} -  b_2  \right) - b_1^2} \le  \frac{ b_1\rho_2 a_2^{-2} }{  \rho_1\rho_2  a_1^{-2}a_2^{-2}-  b_0\rho_1 a_1^{-2} - b_2\rho_2  a_2^{-2}} \nonumber\\
%&\le  \frac{ \rho_2 (b_1a_1^{2}) }{  \rho_1\rho_2 -  \rho_1  - \rho_2 } \le \frac{\rho_2\cdot \mu_{\min}^{-2}}{\rho_1\rho_2-\rho_1-\rho_2}=\OO\left( \frac{1}{\rho_1}\right),\label{est_a4extra}
%\end{align}
%where in the above derivations we also used that $b_1^2\le b_0 b_2$ by Cauchy-Schwarz inequality, $b_0 a_2^2 < 1$, $b_2 a_1^2 < 1$, and $b_1 a_1^2 \le \mu_{\min}^{-2}$.
%Inserting \eqref{est_a12extra} into \eqref{gvar_extra}, we obtain that with high probability,
%\begin{align}\label{approximateteM}
%\le g(M_0) \le  \frac{\sigma^2}{\rho_1+\rho_2}\cdot \frac1p\tr\left( \frac{1}{a_1M_0^\top M_0}\right) \cdot \left(1 +\OO(p^{-\e})\right)
%\end{align}
%where
%$$\wt g(M_0):=\frac{\sigma^2}{\rho_1+\rho_2}\cdot \frac1p\tr\left( \frac{1}{a_1M_0^\top M_0}\right)  +\frac{pd^2\cdot \rho_1^2}{(\rho_1 + \rho_2)^2} \cdot \frac1p\tr\left(\Sigma_1\frac{1 + a_3}{a_1^2 M_0^{\top}{M_0}} \right),$$
%and the error satisfies
%$$|\cal E|\le C\left( \frac{\rho_2}{\rho_1} +\rho_1^{-1/2} +\OO(p^{-\e})\right)\wt g(M_0).$$
%Here the constant $C>0$ depends only on $\mu_{\max}$, $\mu_{\min}$ and $\|\Sigma_1\|$, but otherwise does not depend on $\rho_1$ and $\rho_2$. 
%--------end old argument---------

\begin{proof}[Theoretical justification of Example \ref{ex_complement}]
We denote the test error as $\te_\lambda(\hat \beta_{t}^{\MTL})$ in the setting where $M$ has $p/2$ singular values that are equal to $\lambda$ and $p/2$ singular values that are equal to $1 / \lambda$. Then equations in \eqref{eq_a12extra} become
\be\label{compleeq} a_1 + a_2 = 1 - \frac{1}{\rho_1 + \rho_2},  \ \ a_1 + \frac{1}{2(\rho_1 + \rho_2)}\cdot \bigbrace{\frac{a_1}{a_1 + \lambda^2 a_2} + \frac{a_1}{a_1 + \frac{a_2}{\lambda^2}}} = \frac{\rho_1}{\rho_1 + \rho_2}. \ee
%It's not hard to verify that there is only one valid solution $(a_1,a_2)$ to \eqref{compleeq}. 
After solving these, with \eqref{gvar_extra} we get that whp,
\be\label{testcomple}
 \te_\lambda(\hat \beta_{t}^{\MTL})= \frac{\sigma^2}{2(\rho_1 + \rho_2)}(1+\OO(p^{-\e}))\cdot g(\lambda) ,\quad g(\lambda): = \frac{1}{\frac{a_1}{\lambda^2} + a_2} + \frac{1}{a_1\lambda^2 + a_2}.\ee

%First we notice that the curves in Figure  \ref{fig_model_shift_phasetrans} (c) all cross at the point $n_1=n_2$. In fact, if $n_1=n_2$, then it is easy to observe that $a_1=a_2=(1-\gamma)/2$ is the solution to equation \eqref{compleeq}, where we denote $ \gamma=p/(n_1+n_2)$. Then for any $\lambda$, the test error in \eqref{testcomple} takes the value
%$$te(\lambda)= \frac{\gamma}{2}\frac{1}{(1-\gamma)/2}=\frac{p}{n_1+n_2-p}.$$

%Second, from Figure \ref{fig_te_complement} we observe that the complementary cases with $\lambda>1$ is better than the case without covariate shift (i.e. $M=\id$ case) when $n_1<n_2$. On the other hand, if we have enough source task data such that $n_1>n_2$, then it is always better to have no covariate shift.
We now study the behavior of $g$ as $\lambda$ changes. 
%This phenomenon can be also explained using our theory. 
We abbreviate $\gamma:=(\rho_1 + \rho_2)^{-1}$. Then with \eqref{compleeq}, we can rewrite
$$g(\lambda)= \frac{1}{\frac{a_1}{\lambda^2} + (1-\gamma - a_1)} + \frac{1}{a_1\lambda^2 + (1-\gamma - a_1)}.$$
We can compute that
\begin{align*}
g(\lambda) - g(1)&= \frac{ \lambda^2-1}{1-\gamma} a_1\cdot \bigbrace{  \frac{1}{ -a_1(\lambda^2-1)+(1-\gamma)\lambda^2 } - \frac{1}{a_1(\lambda^2-1) + (1-\gamma)}} \\
&= \frac{(\lambda^2-1)^2}{1-\gamma}  a_1\cdot  \frac{2a_1 - (1-\gamma) }{[-a_1(\lambda^2-1)+(1-\gamma)\lambda^2 ][a_1(\lambda^2-1) + (1-\gamma)]} .
\end{align*}
From this expressions, we observe the following behaviors.
\begin{itemize}
\item[(i)] If $n_1>n_2$, we have $a_1>(1-\gamma)/2$ (because $a_1>a_2$ as observed from the equation \eqref{compleeq}). Hence $g(\lambda)>g(1)$, which gives $\te_\lambda(\hat \beta_{t}^{\MTL})>\te_1(\hat \beta_{t}^{\MTL})$.

\item[(ii)] If $n_1< n_2$, we have $a_1< (1-\gamma)/2$. Hence $g(\lambda)< g(1)$, which gives $\te_\lambda(\hat \beta_{t}^{\MTL})<\te_1(\hat \beta_{t}^{\MTL})$.

\item[(iii)] If $n_1=n_2$, we have $g(\lambda)=g(1)=2/(1-\gamma)$, which explains why the curves in Figure  \ref{fig_model_shift_phasetrans} (c) all cross at the point $n_1=n_2$.
\end{itemize}
These justify the observations in Figure \ref{fig_model_shift_phasetrans} (c), 
\end{proof}
\todo{add a transition example}

@@

\subsection{Proofs for Section \ref{sec_benefit}}\label{app_proof_data}

 We first prove Proposition \ref{prop_data_efficiency}, which gives precise upper and lower bounds on the data efficiency ratio.

\begin{proof}[Proof of Proposition \ref{prop_data_efficiency}]
Suppose we have reduced number of datapoints---$\al n_1$ for task 1 and $\al n_2$ for task 2 with $n_1=\rho_1 p$ and $n_2=\rho_2 p$. Then all the results in the proof of Proposition \ref{prop_dist_transition} still hold, except that we need to replace $(\rho_1,\rho_2)$ with $(\al\rho_1,\al\rho_2)$. More precisely, we have
	\begin{align*}
		 a_1 = \frac{\rho_1(\al\rho_1 + \al\rho_2 - 1)}{\al(\rho_1 + \rho_2)^2} ,\quad
		& a_2 = \frac{\rho_2(\al\rho_1 + \al\rho_2 - 1)}{\al(\rho_1 + \rho_2)^2} ,  \\
		 a_3 = \frac{\rho_2}{(\rho_1 + \rho_2)(\al\rho_1 + \al\rho_2 - 1)}, \quad
		& a_4 = \frac{\rho_1}{(\rho_1 + \rho_2)(\al\rho_1 + \al\rho_2 - 1)}. 
	\end{align*}
Moreover, we have that with high probability,
	\be\label{reduceproof1}
\begin{split}
\te_i(\hat \beta(\al)) =  \frac{\sigma^2}{\al (\rho_1+\rho_2) - 1}\left(1+ \oo(1)\right)+ \delta^{(i)}_{\beta}, \quad i=1,2.
\end{split}
\ee
Here the model shift biases $\delta^{(i)}_{\beta}$ satisfy that  
\be\nonumber %\label{Deltabeta} 
\al_-^2(\al\rho_i) - \oo(1)  \le {\delta_\beta^{(i)}}/{ \Delta_{\beta}^{(i)}} \le \al_+^2(\al\rho_i) +  \oo(1) , \quad i=1,2,\ee	
%In the setting of Proposition \ref{prop_data_efficiency}, we still have \eqref{hatw_add1}. 
%%and by symmetry it suffices to focus on one of tasks, say task 2. 
%Then using Lemmas \ref{lem_cov_shift} and \ref{prop_model_shift_tight}, and recalling \eqref{simplesovlea12} and \eqref{simplesovlea34}, we get that whp,
%\be\label{reduceproof1}
%\begin{split}
%\te_i(\hat \beta(\al)) =\sigma^2\left( \frac{1}{\al (\rho_1+\rho_2) - 1}+ \bigo{p^{-1/2 + \varepsilon}}\right)+ \Delta^{(i)}_{\beta}, \quad i=1,2,
%\end{split}
%\ee
%and
%\be\label{reduceproof1}
%\begin{split}
%\te_2(\hat \beta(\al)) =\sigma^2\left( \frac{1}{\al ( \rho_1+\rho_2) - 1}+ \Delta^{(2)}_{\beta}+ \bigo{p^{-1/2 + \varepsilon}}\right),
%\end{split}
%\ee
where $ \Delta^{(i)}_{\beta}$ are defined as 
\be \nonumber
\Delta^{(i)}_{\beta} := pd^2 \frac{(\al\rho_i)^2\cdot \al (\rho_1+\rho_2)}{[\al ( \rho_1+\rho_2) - 1]^3} ,\quad i=1,2,. 
\ee
On the other hand, using Lemma \ref{lem_minv}, we have whp,
\be\label{reduceproof2} 
\te_i(\hat{\beta}_t^{\STL}) = \frac{\sigma^2}{\rho_i -1} \left( 1+ \oo(1)\right),\quad i=1,2.
\ee
Comparing \eqref{reduceproof1} and \eqref{reduceproof2}, we immediately obtain a trivial lower bound $\al^*\ge \al_l -\oo(1)$, where
$$\al_l:=\frac1{\rho_1+\rho_2}\left(\frac{2}{\frac1{\rho_1-1}+\frac1{\rho_2 -1}}+1\right) \ge \frac{\min(\rho_1,\rho_2)}{\rho_1+\rho_2}.$$
In fact, one can see that if $\al\le \al_l$, then we have 
$$ \frac{2\sigma^2}{\al (\rho_1+\rho_2) - 1} \ge \frac{\sigma^2}{\rho_1-1}+\frac{\sigma^2}{\rho_2-1},$$
that is $\te_1(\hat{\beta}(\alpha)) + \te_2(\hat{\beta}(\alpha))$ is larger than $\te_1(\hat{\beta}_t^{\STL}) + \te_2(\hat{\beta}_t^{\STL})$ even if we do not take into account the model shift bias terms $ \delta^{(i)}_{\beta}$. Next we try to get more precise bounds on $\al^*$. In the following discussions, we only consider $\al$ such that $\al(\rho_1+\rho_2) \ge \al_l (\rho_1+\rho_2) \ge \min(\rho_1,\rho_2)$.
